Category-Level Transfer Learning from Knowledge Base to Microblog Stream for Accurate Event Detection

### 2016-12-26 paper/lncsFormat.tex
TODO: camera-ready

\usepackage{hyperref}
\setcounter{secnumdepth}{2}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Category-Level Transfer Learning from Knowledge Base to Microblog Stream for Accurate Event Detection}
%\title{Leveraging Knowledge Base for Accurate Event Detection on Microblog Stream}
\titlerunning{\ }
%构造（building）一个**词表（找一个词对它更合适的描述），for transfer learning in accurate event detection
%\title{Constructing Knowledge Base's Category Topics for Transfer Learning in Accurate Microblog Event Detection}
\begin{comment}% for double blind
\author{Weijing Huang\inst{1} \and Wei Chen\inst{1} \thanks{Corresponding author.} \and Tengjiao Wang\inst{1} \and Yazhou Wang\inst{1} }
\institute{Key Laboratory of High Confidence Software Technologies (Ministry of Education), EECS, Peking University, Beijing, China \\ 
\path{huangwaleking@gmail.com,{pekingchenwei,tjwang}@pku.edu.cn}\\ 
\path{pkuwangyz@gmail.com}
}
\end{comment}

@@

Knowledge base can be a good supplementary for event detection on microblog stream to address these challenges.
Different from the not-well-organized microblog stream, knowledge base (e.g. Wikipedia) is constructed elaborately and contains rich information. 
For example, the microblog message ``Possible Ft. Hood Attack Thwarted (2011-07-28)'' is short, but still comprehensible because the words ``\textit{Ft. Hood}" is included in the wiki page ``\textit{Ft. Hood}", and belongs to the category ``\textit{Military}'' at a higher conceptual level.
By regarding these wiki information and the word \textit{attack} (also highly related to \textit{Military}), the model easily understands the example tweet is about something related to \textit{Military}.
In other words, knowledge base enriches the linkages between words and concepts, and provides more comprehensive context for microblogs.
Since the transfer learning\cite{pan2010survey} aims at utilizing the extra information restored in the source dataset to benefit the target dataset, it provides a feasible way to enhance the event detection in microblog stream.
Taking the below Figure \ref{fig:hood} as an example, the fluctuation of time series of raw word \textit{hood} is not so obvious to reflect the event happened to the military base \textit{Ft. Hood}.
After transferring the knowledge about \textit{``Military"} into the microblog stream, the time series of word \textit{``hood"} related to \textit{``Military"} is extracted, which is more vivid to detect the event happened on July 28th, 2011.

@@

The benefit of enriching the semantics for micorblogs is attractive, but it's non-trivial to transfer the knowledge restored in the knowledge base into microblog stream directly.
The existing RDF model\cite{klyne2006rdf} lacks an efficient quick mechanism to transfer the knowledge, since it's mainly designed for managing knowledge as tuples on graph. And the query on large graph is also very expensive\cite{huang2011scalable}, which is not suitable for the scene of quickly and accurately detecting events.
In existing methods, what meets the demand most is Twevent\cite{Twevent2012}, but it's limited in only treating Wikipedia as a looking-up table and may drop some events incorrectly.

In our paper, to balance the performance and the cost of leveraging knowledge base for event detection, we develop a novel category-level transfer learning method, namely \textsc{TransDetector}.
It consists of three phases: extracting category-level topics in knowledge base, conducting transfer learning to get category-level topics in microblog stream, and detecting events from category-level word time series, as illustrated in Figure \ref{fig:modelDesc}.
We explain the main idea of the three parts in \textsc{TransDetector}, and leave more technique details in Section \ref{sec:TransDetector}.
(1) In the \textbf{extracting} phase, we propose a structural-guided category-level topics extraction method on the knowledge base, . 
It bases on the following facts. 
The knowledge base has the three fold hierarchical structure, consisting of the taxonomy graph (\textit{class} \(\rightarrow\) \textit{subclass}), the category-page bipartite graph (\textit{class} \(\rightarrow\) \textit{instance}), and the page-content map (\textit{instance} \(\rightarrow\) \textit{content}), as illustrated in Figure \ref{fig:NSinitializaton}.
In terms of concept level, the latter part is finer than the former. 
And the last page-content map goes into the detail at the word level. 
By considering these three parts together, we can extract a class or category's highly related words, and restore them in \textit{category-level topics in knowledge base}.
(2) In the \textbf{conducting transfer learning} phase, we propose a novel probabilistic model CTrans-LDA for transferring the knowledge.
CTrans-LDA works in the bayesian transfer learning way like \cite{dai2007transferring}, by utilizing the extracted \textit{category-level topics in knowledge base} as the informative priors to bridge two data domains. 
CTrans-LDA labels whether a word in a microblog message should link to a category in knowledge base, or just label it as no-category-related word.
Applying CTrans-LDA on more microblogs, it gets the \textit{category-level topics in microblog stream} and \textit{category-level word time series}.
(3) In the \textbf{detecting} phase, since the words in microblogs have been enriched semantically and the meaningless words are labeled as no-category-related, the event detection on \textit{category-level word time series} is much more accurate than other methods which don't conduct the transfer learning.
Without loss of generality, we use Wikipedia as a representative of the knowledge base in the remaining of this paper.
And the experiment on the \textit{Edinburgh twitter corpus} demonstrates the effectiveness of our proposed \textsc{TransDetector}.

@@

\section{Related Works}
\label{sec:relatedWorks}
\textbf{Event Detection.} Based on how much data used, the methods are mainly in two groups, the detection \textit{without extra information} and that by \textit{leveraging extra information}.

@@

\textbf{Page-Content Map \(G^{(2)}\)}. 
For a specific Wikipedia dumps version, the edges  \textit{page} \(\rightarrow\)\textit{content} in \(G^{(2)}\) define a one-to-one mapping.
There are a bulk of information restored in the wiki text content.
In order to extract the key words in wiki page's content, which distinguish it from other pages, we use the chi-square statistics\cite{yang1997comparative}\cite{liu2009imbalanced} to measure the importance of each word for the specific page.

@@

For CTrans-LDA, \(K\) is set to be 200, which means CTrans-LDA learned 50 category-level topics and 150 other topics.
After cross-validated grid-search, \(\lambda\) is set to be 12.8. The other parameters are set as \(\alpha=0.1\), \(\beta=0.005\).
We run CTrans-LDA window by window on text stream, and learn specific categories' topics.
The result in Table 2 demonstrates the effectiveness of transferring category-level topics into the microblog stream, and finding more new words in the stream which is not restored in the knowledge base.

>>>>

\usepackage{hyperref}
\setcounter{secnumdepth}{2}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Category-Level Transfer Learning from Knowledge Base to Microblog Stream for Accurate Event Detection}
\titlerunning{\ }
\author{Weijing Huang\inst{1} \and Wei Chen\inst{1} \thanks{Corresponding author.} \and Tengjiao Wang\inst{1} \and Yazhou Wang\inst{1} }
\institute{Key Laboratory of High Confidence Software Technologies (Ministry of Education), EECS, Peking University, Beijing, China \\ 
\path{huangwaleking@gmail.com,{pekingchenwei,tjwang}@pku.edu.cn}\\ 
\path{pkuwangyz@gmail.com}
}

@@

Knowledge base can be a good supplementary for event detection on microblog stream to address these challenges.
Different from the not-well-organized microblog stream, knowledge base (e.g. Wikipedia) is constructed elaborately and contains rich information. 
For example, the microblog message ``Possible Ft. Hood Attack Thwarted (2011-07-28)'' is short, but still comprehensible because the words ``\textit{Ft. Hood}" is included in the wiki page ``\textit{Ft. Hood}", and belongs to the category ``\textit{Military}'' at a higher conceptual level.
By regarding these wiki information and the word \textit{attack} (also highly related to \textit{Military}), the model easily understands the example tweet is about something related to \textit{Military}.
In other words, knowledge base enriches the linkages between words and concepts, and provides more comprehensive context for microblogs.
Since the transfer learning\cite{pan2010survey} aims at utilizing the extra information stored in the source dataset to benefit the target dataset, it provides a feasible way to enhance the event detection in microblog stream.
Taking the below Figure \ref{fig:hood} as an example, the fluctuation of time series of raw word \textit{hood} is not so obvious to reflect the event happened to the military base \textit{Ft. Hood}.
After transferring the knowledge about \textit{``Military"} into the microblog stream, the time series of word \textit{``hood"} related to \textit{``Military"} is extracted, which is more vivid to detect the event happened on July 28th, 2011.

@@

The benefit of enriching the semantics for micorblogs is attractive, but it's non-trivial to transfer the knowledge stored in the knowledge base into microblog stream directly.
The existing RDF model\cite{klyne2006rdf} lacks an efficient quick mechanism to transfer the knowledge, since it's mainly designed for managing knowledge as tuples on graph. And the query on large graph is also very expensive\cite{huang2011scalable}, which is not suitable for the scene of quickly and accurately detecting events.
In existing methods, what meets the demand most is Twevent\cite{Twevent2012}, but it's limited in only treating Wikipedia as a looking-up table and may drop some events incorrectly.

In our paper, to balance the performance and the cost of leveraging knowledge base for event detection, we develop a novel category-level transfer learning method, namely \textsc{TransDetector}.
It consists of three phases: extracting category-level topics in knowledge base, conducting transfer learning to get category-level topics in microblog stream, and detecting events from category-level word time series, as illustrated in Figure \ref{fig:modelDesc}.
We explain the main idea of the three parts in \textsc{TransDetector}, and leave more technique details in Section \ref{sec:TransDetector}.
(1) In the \textbf{extracting} phase, we propose a structural-guided category-level topics extraction method on the knowledge base. 
It bases on the following facts. 
The knowledge base has the three fold hierarchical structure, consisting of the taxonomy graph (\textit{class} \(\rightarrow\) \textit{subclass}), the category-page bipartite graph (\textit{class} \(\rightarrow\) \textit{instance}), and the page-content map (\textit{instance} \(\rightarrow\) \textit{content}), as illustrated in Figure \ref{fig:NSinitializaton}.
In terms of concept level, the latter part is finer than the former. 
And the last page-content map goes into the detail at the word level. 
By considering these three parts together, we can extract a class or category's highly related words, and restore them in \textit{category-level topics in knowledge base}.
(2) In the \textbf{conducting transfer learning} phase, we propose a novel probabilistic model CTrans-LDA for transferring the knowledge.
CTrans-LDA works in the bayesian transfer learning way like \cite{dai2007transferring}, by utilizing the extracted \textit{category-level topics in knowledge base} as the informative priors to bridge two data domains. 
CTrans-LDA labels whether a word in a microblog message should link to a category in knowledge base, or just label it as no-category-related word.
Applying CTrans-LDA on more microblogs, it gets the \textit{category-level topics in microblog stream} and \textit{category-level word time series}.
(3) In the \textbf{detecting} phase, since the words in microblogs have been enriched semantically and the meaningless words are labeled as no-category-related, the event detection on \textit{category-level word time series} is much more accurate than other methods which don't conduct the transfer learning.
Without loss of generality, we use Wikipedia as a representative of the knowledge base in the remaining of this paper.
And the experiment on the \textit{Edinburgh twitter corpus} demonstrates the effectiveness of our proposed \textsc{TransDetector}.

@@

\section{Related Works}
\label{sec:relatedWorks}
\textbf{Event Detection.} Based on how much data are used, the methods are mainly in two groups, the detection \textit{without extra information} and that by \textit{leveraging extra information}.

@@

\textbf{Page-Content Map \(G^{(2)}\)}. 
For a specific Wikipedia dumps version, the edges  \textit{page} \(\rightarrow\)\textit{content} in \(G^{(2)}\) define a one-to-one mapping.
There are a bulk of information stored in the wiki text content.
In order to extract the key words in wiki page's content, which distinguish it from other pages, we use the chi-square statistics\cite{yang1997comparative}\cite{liu2009imbalanced} to measure the importance of each word for the specific page.

@@

For CTrans-LDA, \(K\) is set to be 200, which means CTrans-LDA learned 50 category-level topics and 150 other topics.
After cross-validated grid-search, \(\lambda\) is set to be 12.8. The other parameters are set as \(\alpha=0.1\), \(\beta=0.005\).
We run CTrans-LDA window by window on text stream, and learn specific categories' topics.
The result in Table 2 demonstrates the effectiveness of transferring category-level topics into the microblog stream, and finding more new words in the stream which is not stored in the knowledge base.
